---
title: "Git and RStudio"
author: "Scott Moser"
date: "2023-05-08"
output: html_document
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "../Outputs"
    )
  })
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## You Will Need:

- R: 
- RStudio
- a GitHub account
- Git on your local computer


## Steps

1. Install git on your local computer <https://sites.northwestern.edu/researchcomputing/resources/using-git-and-github-with-r-rstudio/> number 2

	1. Make sure R: knows where `git` is located: <https://sites.northwestern.edu/researchcomputing/resources/using-git-and-github-with-r-rstudio/>
	2. Configure `git` from command line: <https://swcarpentry.github.io/git-novice/02-setup/index.html>  and `git config --system core.longpaths true
`
	
2. Get Personal Authorization Token (PAT) from <https://github.com/settings/tokens>: <https://happygitwithr.com/https-pat.html>

	1. Let `git` know about your PAT <https://happygitwithr.com/https-pat.html>.  In R:
	```{r eval=FALSE}
	## install.packages("gitcreds")
	gitcreds::gitcreds_set()
	```
	And paste token.

3. Create local RProj based on clone from GitHub <https://rfortherestofus.com/2021/02/how-to-use-git-github-with-r/>   GitHub First section

	1. <https://github.com/smoser11/ArmedPeaceData.git>
	2. Take note of local file structure

```{r}
 
```
	
	
	
## Git Resources

- Workflows <https://happygitwithr.com/workflows-intro.html>
- Git commands <https://happygitwithr.com/git-intro.html>
-Staging, Committing, Push/Pull
